Lineage Analysis of CRISPR and shRNA Screens

### 2018-10-16 lineages_18Q3.Rmd

Load libraries.
```{r}
library(broom)
library(bsselectR)
library(caret)
library(CePa)
library(circlize)
library(ComplexHeatmap)
library(data.table)
library(devtools)
library(ggpubr)
library(ggsignif)
library(ggfortify)
library(ggord)
library(glmnet)
library(gridExtra)
# devtools::install_github("jeroen/jsonlite")
# install.packages("jsonlite")
library(jsonlite)
library(jsonlite)
library(kableExtra)
library(magrittr)
library(matrixStats)
library(NMF)
library(parallel)
library(plyr)
library(reshape2)
library(rowr)
library(tidyverse)
```

@@

## Make CRISPR grobs

@@

## Make lineage plot

@@

# Data management

@@

# Select columns
maf_df <- subset(maf_raw, select = c("Hugo_Symbol", "CCLE_Name", "Broad_ID", "Variant_Classification", "isDeleterious", "Reference_Allele", "Tumor_Seq_Allele1"))
maf_df$Reference_Allele_Length <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1_Length <- nchar(as.character(maf_df$Tumor_Seq_Allele1))

@@

```{r, eval=FALSE}
max(nchar(as.character(maf_raw$Reference_Allele)))
maf_raw %>% group_by(Variant_Type, isTCGAhotspot) %>% tally()

@@

```{r}
cn_melt <- readRDS("./data_munging/rds/cn_melt_18Q3.rds")
```

@@

```{r}
ge_melt <- readRDS("./../crispr_lineages_giant_files/ge_melt_18Q3.rds")
```

@@

```{r fig.height=12, fig.width=12, eval=FALSE}
crispr_pca_data <- merge(ccl_converter[, c("Broad_ID", "group_general_lineage_name")], crispr, by = "Broad_ID")
crispr_lin <- crispr_pca_data[, 2]
crispr_pca <- prcomp(data.matrix(crispr_pca_data[, 3:ncol(crispr_pca_data)]), scale. = TRUE, center = TRUE)
plot(crispr_pca, type = "l")
biplot(crispr_pca, obs.scale = 1, var.scale = 1, groups = crispr_lin, ellipse = TRUE, circle = TRUE)
ggord(crispr_pca, crispr_pca_data$group_general_lineage_name)
```

Merge annotation data:
```{r, eval=FALSE}
# Melt CRISPR dataset for merging
crispr_melt <- melt(crispr, id.vars = "Broad_ID", measure.vars = colnames(crispr)[2:ncol(crispr)], variable.name = "Hugo_Symbol", value.name = "Score")

# Melt dependency probabilities dataset for merging
dep_melt <- melt(dep, id.vars = "Broad_ID", measure.vars = colnames(dep)[2:ncol(dep)], variable.name = "Hugo_Symbol", value.name = "Dep_Prob")

# Merge dependency probabilities
crispr_melt <- merge(crispr_melt, dep_melt, by = c("Broad_ID", "Hugo_Symbol"), all.x = TRUE)

# Merge cell line metadata
crispr_melt <- merge(crispr_melt, ccl_info, by = "Broad_ID", all.x = TRUE)
crispr_melt <- merge(crispr_melt, crispr_meta, by = c("CCLE_Name", "Broad_ID"), all.x = TRUE)

# Merge mutation annotations
crispr_muts <- merge(crispr_melt, maf_df, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"), all.x = TRUE)
crispr_muts$Hugo_Symbol <- factor(crispr_muts$Hugo_Symbol)
crispr_muts <- crispr_muts %>% mutate(Mutation_Status_Nonsilent = if_else(is.na(Mutation_Status_Nonsilent), "Other", Mutation_Status_Nonsilent))

# Summarize number of mutant and Other cell lines
crispr_muts_summ <- crispr_muts %>% group_by(Hugo_Symbol) %>%
  summarize(N_Nonsilent_Other = sum(Mutation_Status_Nonsilent == "Other"),
            N_Nonsilent_Mutant = sum(Mutation_Status_Nonsilent == "Mutant"))

# Merge test results back into full dataset, which restores information lost in the summarization
crispr_data <- merge(crispr_muts_summ, crispr_muts, by = "Hugo_Symbol")

# Add Color columns
crispr_data$Color_Nonsilent <- ifelse(crispr_data$Mutation_Status_Nonsilent == "Other", "thistle", "darkorchid")
crispr_data$Color_Nonsilent <- factor(crispr_data$Color_Nonsilent)

# Cell line lineages
crispr_data <- merge(crispr_data, ccl_converter, by = c("CCLE_Name", "Broad_ID"), all.x = TRUE)
levels(crispr_data$lineage_name) <- sort(levels(crispr_data$lineage_name), decreasing = TRUE)

# Copy number
crispr_data <- merge(crispr_data, cn_melt, by = c("Hugo_Symbol", "Broad_ID"), all.x = TRUE)

# Gene expression (RPKM)
ge_filt <- filter(ge_melt, Hugo_Symbol %in% unique(crispr_data$Hugo_Symbol) & Broad_ID %in% unique(crispr_data$Broad_ID))
crispr_data <- merge(crispr_data, ge_filt, by = c("Hugo_Symbol", "Broad_ID", "CCLE_Name"), all.x = TRUE)
crispr_data$RPKM_log2 <- log2(crispr_data$RPKM + 0.0001)

saveRDS(crispr_data, "./../crispr_lineages_giant_files/crispr_data_18Q3.rds", compress = "xz")
```

@@

@@

```{r, eval=FALSE}
# Drug metadata
drug_meta <- read.delim("./data_munging/drug_id_drug_name_table_10_12_2018.tsv", sep = "\t", header = FALSE)
colnames(drug_meta) <- c("CID", "Drug")
drug_meta$Drug <- tolower(drug_meta$Drug)

# CCLE
ccle <- read.delim("./data_munging/data_drug_auc_ccle_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ccle <- merge(drug_meta, ccle, by = "CID")
ccle_melt <- melt(ccle, id.vars = c("CID", "Drug"), measure.vars = colnames(ccle)[3:ncol(ccle)], variable.name = "accession_id", value.name = "AUC_CCLE")

# CTRP
ctrp <- read.delim("./data_munging/data_drug_auc_ctrp_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ctrp <- merge(drug_meta, ctrp, by = "CID")
ctrp_melt <- melt(ctrp, id.vars = c("CID", "Drug"), measure.vars = colnames(ctrp)[3:ncol(ctrp)], variable.name = "accession_id", value.name = "AUC_CTRP")

# GDSC
gdsc <- read.delim("./data_munging/data_drug_auc_gdsc_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
gdsc <- merge(drug_meta, gdsc, by = "CID")
gdsc_melt <- melt(gdsc, id.vars = c("CID", "Drug"), measure.vars = colnames(gdsc)[3:ncol(gdsc)], variable.name = "accession_id", value.name = "AUC_GDSC")

# Merge all data
drug_data <- merge(merge(ccle_melt, ctrp_melt, by = c("accession_id", "CID", "Drug"), all = TRUE), gdsc_melt, by = c("accession_id", "CID", "Drug"), all = TRUE)
drug_data <- merge(drug_data, ccl_converter, by = "accession_id", all.x = TRUE)
drug_data <- merge(drug_data, maf_df, by = c("CCLE_Name", "Broad_ID"))
drug_data$Hugo_Symbol <- factor(drug_data$Hugo_Symbol)

saveRDS(drug_data, "./data_munging/drug_data_18Q3.rds", compress = "xz")
```

```{r}
drug_data <- readRDS("./data_munging/drug_data_18Q3.rds")
```

# Wilcoxon tests

@@

Filter for point mutations:
```{r}
crispr_data_ptmuts <- filter(crispr_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))

crispr_data_ptmuts$Keep_PtMut_Tests <- ifelse(crispr_data_ptmuts$Reference_Allele_Length == 1 | is.na(crispr_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(crispr_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))

crispr_data_ptmuts <- filter(crispr_data_ptmuts, Keep_PtMut_Tests == TRUE)

@@

# Paper outline: G2P data

@@

```{r}
g2p <- read.delim("./data_munging/g2p_full_level_A_dataset_v2_10_10_2018.tsv", sep = "\t", header = TRUE)
genes_lvlA <- unique(g2p$Gene)
```

@@

### Gene
```{r, eval=FALSE}
crispr_signif_lvlA <- compare_means(Score ~ Mutation_Status_Nonsilent, group.by = c("Hugo_Symbol"), data = crispr_data_lvlA, method = "wilcox.test", p.adjust.method = "BH")
crispr_signif_lvlA <- adj_signif(crispr_signif_lvlA)
crispr_signif_lvlA <- crispr_signif_lvlA[order(crispr_signif_lvlA$p),]
saveRDS(crispr_signif_lvlA, "./data_munging/rds/crispr_signif_lvlA_gene.rds")
```

@@

crispr_signif_genes_lvlA <- as.character(filter(crispr_signif_lvlA, !is.na(p.signif.adj))$Hugo_Symbol)

knitr::kable(filter(crispr_signif_lvlA, p < 0.05)[, c("Hugo_Symbol", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "Wilcoxon test results for Level A point mutations, p < 0.01 (BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

@@

### Gene and lineage

@@

## Figure 1C plots

Mutation status boxplots for genes identified in G2P.

```{r fig.height=4, fig.width=20}
genes_lvlA_filt <- c("ABL1", "ALK", "AR", "ATM", "BRAF", "BRCA1", "BRCA2", "CDKN2A", "DNMT3A", "DPYD", "EGFR", "ERBB2", "ERBB3", "FGFR2", "FLT3", "JAK2", "KIT", "KRAS", "MET", "NOTCH1", "NOTCH2", "NRAS", "PDGFRA", "PDGFRB", "PIK3CA", "PML", "PTCH1", "RET", "ROS1", "SF3B1", "SMO", "TP53", "TSC1")
crispr_ptmuts_lvlA_filt <- filter(crispr_data_ptmuts, Hugo_Symbol %in% genes_lvlA_filt)
crispr_ptmuts_lvlA_filt$Hugo_Symbol <- ordered(crispr_ptmuts_lvlA_filt$Hugo_Symbol, levels = genes_lvlA_filt)
crispr_ptmuts_lvlA_filt_color <- as.character(crispr_ptmuts_lvlA_filt$Color_Nonsilent)
names(crispr_ptmuts_lvlA_filt_color) <- crispr_ptmuts_lvlA_filt$Mutation_Status_Nonsilent

>>>>

Load libraries.
```{r}
library(broom)
library(bsselectR)
library(caret)
library(CePa)
library(circlize)
library(ComplexHeatmap)
library(data.table)
library(devtools)
library(ggpubr)
library(ggsignif)
library(ggfortify)
library(ggord)
library(glmnet)
library(gridExtra)
library(jsonlite)
library(jsonlite)
library(kableExtra)
library(magrittr)
library(matrixStats)
library(NMF)
library(parallel)
library(plyr)
library(reshape2)
library(rowr)
library(tidyverse)
```

@@

## Make CRISPR grouped plots

@@

## Make sorted lineage plots

@@

## Make drug screen boxplots
```{r}
makeDrugBoxplots <- function(drug, dataset) {
  if(dataset == "CCLE" | dataset == "ccle") {
    use_data <- filter(ccle_ptmuts_lvlA_filt, Drug == drug)
    use_color <- as.character(use_data$Color_Nonsilent)
    names(use_color) <- use_data$Mutation_Status_Nonsilent
    sig <- filter(ccle_signif_lvlA, Drug == drug)
    label_text <- data.frame(p.signif.adj = sig$p.signif.adj, Hugo_Symbol = sig$Hugo_Symbol)
    label_text <- filter(label_text, Hugo_Symbol %in% genes_lvlA_filt)
    label_text$Hugo_Symbol <- factor(label_text$Hugo_Symbol, levels = genes_lvlA_filt)
  }
  else if(dataset == "CTRP" | dataset == "ctrp") {
    use_data <- filter(ctrp_ptmuts_lvlA_filt, Drug == drug)
    use_color <- as.character(use_data$Color_Nonsilent)
    names(use_color) <- use_data$Mutation_Status_Nonsilent
    sig <- filter(ctrp_signif_lvlA, Drug == drug)
    label_text <- data.frame(p.signif.adj = sig$p.signif.adj, Hugo_Symbol = sig$Hugo_Symbol)
    label_text <- filter(label_text, Hugo_Symbol %in% genes_lvlA_filt)
    label_text$Hugo_Symbol <- factor(label_text$Hugo_Symbol, levels = genes_lvlA_filt)
  }
  else if(dataset == "GDSC" | dataset == "gdsc") {
    use_data <- filter(gdsc_ptmuts_lvlA_filt, Drug == drug)
    use_color <- as.character(use_data$Color_Nonsilent)
    names(use_color) <- use_data$Mutation_Status_Nonsilent
    sig <- filter(gdsc_signif_lvlA, Drug == drug)
    label_text <- data.frame(p.signif.adj = sig$p.signif.adj, Hugo_Symbol = sig$Hugo_Symbol)
    label_text <- filter(label_text, Hugo_Symbol %in% genes_lvlA_filt)
    label_text$Hugo_Symbol <- factor(label_text$Hugo_Symbol, levels = genes_lvlA_filt)
  }
  else { return("Error: Invalid dataset.") }

  plot <- ggplot(data = use_data, aes(x = Mutation_Status_Nonsilent, y = AUC)) +
    facet_wrap(~ Hugo_Symbol, drop = FALSE, ncol = length(genes_lvlA_filt)) +
    geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
    geom_boxplot(mapping = aes(fill = Mutation_Status_Nonsilent), position = position_dodge(0.85)) +
    scale_fill_manual(values = use_color) +
    guides(color = FALSE) +
    geom_text(data = label_text, mapping = aes(x = 1.5, y = max(use_data$AUC), label = p.signif.adj)) +
    theme(legend.position = "top", axis.ticks.x = element_blank(), axis.text.x = element_blank(), axis.title.x = element_blank()) +
    labs(fill = "Mutation Status", y = "AUC", x = "Mutation Status", title = paste0(dataset, ": ", drug))
  return(plot)
}
```

# Data management {.tabset .tabset-fade .tabset-pills}

@@

# Select columns
maf_df <- subset(maf_raw, select = c("Hugo_Symbol", "CCLE_Name", "Broad_ID", "Variant_Classification", "Reference_Allele", "Tumor_Seq_Allele1"))
maf_df$Reference_Allele_Length <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1_Length <- nchar(as.character(maf_df$Tumor_Seq_Allele1))

@@

```{r, eval=FALSE, include=FALSE}
max(nchar(as.character(maf_raw$Reference_Allele)))
maf_raw %>% group_by(Variant_Type, isTCGAhotspot) %>% tally()

@@

```{r, eval=FALSE}
cn_melt <- readRDS("./data_munging/rds/cn_melt_18Q3.rds")
```

@@

```{r, eval=FALSE}
ge_melt <- readRDS("./../crispr_lineages_giant_files/ge_melt_18Q3.rds")
```

@@

Merge annotation data:
```{r, eval=FALSE}
# Melt CRISPR dataset for merging
crispr_melt <- melt(crispr, id.vars = "Broad_ID", measure.vars = colnames(crispr)[2:ncol(crispr)], variable.name = "Hugo_Symbol", value.name = "Score")
# Melt dependency probabilities dataset for merging
dep_melt <- melt(dep, id.vars = "Broad_ID", measure.vars = colnames(dep)[2:ncol(dep)], variable.name = "Hugo_Symbol", value.name = "Dep_Prob")
# Merge dependency probabilities
crispr_melt <- merge(crispr_melt, dep_melt, by = c("Broad_ID", "Hugo_Symbol"), all.x = TRUE)
# Merge cell line metadata
crispr_melt <- merge(crispr_melt, ccl_info, by = "Broad_ID", all.x = TRUE)
crispr_melt <- merge(crispr_melt, crispr_meta, by = c("CCLE_Name", "Broad_ID"), all.x = TRUE)
# Merge mutation annotations
crispr_muts <- merge(crispr_melt, maf_df, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"), all.x = TRUE)
crispr_muts$Hugo_Symbol <- factor(crispr_muts$Hugo_Symbol)
crispr_muts <- crispr_muts %>% mutate(Mutation_Status_Nonsilent = if_else(is.na(Mutation_Status_Nonsilent), "Other", Mutation_Status_Nonsilent))
# Summarize number of mutant and Other cell lines
crispr_muts_summ <- crispr_muts %>% group_by(Hugo_Symbol) %>%
  summarize(N_Nonsilent_Other = sum(Mutation_Status_Nonsilent == "Other"),
            N_Nonsilent_Mutant = sum(Mutation_Status_Nonsilent == "Mutant"))
# Merge test results back into full dataset, which restores information lost in the summarization
crispr_data <- merge(crispr_muts_summ, crispr_muts, by = "Hugo_Symbol")
# Add Color columns
crispr_data$Color_Nonsilent <- ifelse(crispr_data$Mutation_Status_Nonsilent == "Other", "thistle", "darkorchid")
crispr_data$Color_Nonsilent <- factor(crispr_data$Color_Nonsilent)
# Cell line lineages
crispr_data <- merge(crispr_data, ccl_converter, by = c("CCLE_Name", "Broad_ID"), all.x = TRUE)
levels(crispr_data$lineage_name) <- sort(levels(crispr_data$lineage_name), decreasing = TRUE)
# Copy number
crispr_data <- merge(crispr_data, cn_melt, by = c("Hugo_Symbol", "Broad_ID"), all.x = TRUE)
# Gene expression (RPKM)
ge_filt <- filter(ge_melt, Hugo_Symbol %in% unique(crispr_data$Hugo_Symbol) & Broad_ID %in% unique(crispr_data$Broad_ID))
crispr_data <- merge(crispr_data, ge_filt, by = c("Hugo_Symbol", "Broad_ID", "CCLE_Name"), all.x = TRUE)
crispr_data$RPKM_log2 <- log2(crispr_data$RPKM + 0.0001)
saveRDS(crispr_data, "./../crispr_lineages_giant_files/crispr_data_18Q3.rds", compress = "xz")
```

@@

```{r fig.height=12, fig.width=12, eval=FALSE, include=FALSE}
# Test dimensional scaling
crispr_pca_data <- merge(ccl_converter[, c("Broad_ID", "group_general_lineage_name")], crispr, by = "Broad_ID")
crispr_lin <- crispr_pca_data[, 2]
crispr_pca <- prcomp(data.matrix(crispr_pca_data[, 3:ncol(crispr_pca_data)]), scale. = TRUE, center = TRUE)
plot(crispr_pca, type = "l")
biplot(crispr_pca, obs.scale = 1, var.scale = 1, groups = crispr_lin, ellipse = TRUE, circle = TRUE)
ggord(crispr_pca, crispr_pca_data$group_general_lineage_name)
```

@@

```{r}
g2p <- read.delim("./data_munging/g2p_full_level_A_dataset_v2_10_10_2018.tsv", sep = "\t", header = TRUE)
genes_lvlA_df <- data.frame(Hugo_Symbol = unique(g2p$Gene))
genes_lvlA <- unique(g2p$Gene)
```

```{r, eval=FALSE}
# Drug metadata
drug_meta <- read.delim("./data_munging/drug_id_drug_name_table_10_12_2018.tsv", sep = "\t", header = FALSE)
colnames(drug_meta) <- c("CID", "Drug")
drug_meta$Drug <- tolower(drug_meta$Drug)
maf_df_lvlA <- filter(maf_df, Hugo_Symbol %in% genes_lvlA)

# CCLE
ccle <- read.delim("./data_munging/data_drug_auc_ccle_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ccle <- merge(drug_meta, ccle, by = "CID")
ccle_melt <- melt(ccle, id.vars = c("CID", "Drug"), measure.vars = colnames(ccle)[3:ncol(ccle)], variable.name = "accession_id", value.name = "AUC")
ccle_grid <- expand.grid("accession_id" = unique(ccle_melt$accession_id), "Hugo_Symbol" = genes_lvlA)
ccle_data <- merge(ccle_grid, ccle_melt, by = "accession_id", all.x = TRUE)
ccle_data <- merge(ccle_data, ccl_converter, by = "accession_id", all.x = TRUE)
ccle_data <- merge(ccle_data, maf_df_lvlA, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"), all.x = TRUE)
ccle_data$Mutation_Status_Nonsilent <- ifelse(is.na(ccle_data$Mutation_Status_Nonsilent), "Other", ccle_data$Mutation_Status_Nonsilent)
ccle_data$Hugo_Symbol <- factor(ccle_data$Hugo_Symbol)
ccle_data$AUC <- as.numeric(ccle_data$AUC)
ccle_data$Color_Nonsilent <- ifelse(ccle_data$Mutation_Status_Nonsilent == "Other", "palegreen3", "springgreen4")
ccle_data$Color_Nonsilent <- factor(ccle_data$Color_Nonsilent)
ccle_data <- filter(ccle_data, !is.na(ccle_data$AUC))
saveRDS(ccle_data, "./data_munging/rds/ccle_data_18Q3.rds", compress = "xz")

# CTRP
ctrp <- read.delim("./data_munging/data_drug_auc_ctrp_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ctrp <- merge(drug_meta, ctrp, by = "CID")
ctrp_melt <- melt(ctrp, id.vars = c("CID", "Drug"), measure.vars = colnames(ctrp)[3:ncol(ctrp)], variable.name = "accession_id", value.name = "AUC")
ctrp_grid <- expand.grid("accession_id" = unique(ctrp_melt$accession_id), "Hugo_Symbol" = genes_lvlA)
ctrp_data <- merge(ctrp_grid, ctrp_melt, by = "accession_id", all.x = TRUE)
ctrp_data <- merge(ctrp_data, ccl_converter, by = "accession_id", all.x = TRUE)
ctrp_data <- merge(ctrp_data, maf_df_lvlA, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"), all.x = TRUE)
ctrp_data$Mutation_Status_Nonsilent <- ifelse(is.na(ctrp_data$Mutation_Status_Nonsilent), "Other", ctrp_data$Mutation_Status_Nonsilent)
ctrp_data$Hugo_Symbol <- factor(ctrp_data$Hugo_Symbol)
ctrp_data$AUC <- as.numeric(ctrp_data$AUC)
ctrp_data$Color_Nonsilent <- ifelse(ctrp_data$Mutation_Status_Nonsilent == "Other", "slategray3", "steelblue4")
ctrp_data$Color_Nonsilent <- factor(ctrp_data$Color_Nonsilent)
ctrp_data <- filter(ctrp_data, !is.na(ctrp_data$AUC))
saveRDS(ctrp_data, "./data_munging/rds/ctrp_data_18Q3.rds", compress = "xz")

# GDSC
gdsc <- read.delim("./data_munging/data_drug_auc_gdsc_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
gdsc <- merge(drug_meta, gdsc, by = "CID")
gdsc_melt <- melt(gdsc, id.vars = c("CID", "Drug"), measure.vars = colnames(gdsc)[3:ncol(gdsc)], variable.name = "accession_id", value.name = "AUC")
gdsc_grid <- expand.grid("accession_id" = unique(gdsc_melt$accession_id), "Hugo_Symbol" = genes_lvlA)
gdsc_data <- merge(gdsc_grid, gdsc_melt, by = "accession_id", all.x = TRUE)
gdsc_data <- merge(gdsc_data, ccl_converter, by = "accession_id", all.x = TRUE)
gdsc_data <- merge(gdsc_data, maf_df_lvlA, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"), all.x = TRUE)
gdsc_data$Mutation_Status_Nonsilent <- ifelse(is.na(gdsc_data$Mutation_Status_Nonsilent), "Other", gdsc_data$Mutation_Status_Nonsilent)
gdsc_data$Hugo_Symbol <- factor(gdsc_data$Hugo_Symbol)
gdsc_data$AUC <- as.numeric(gdsc_data$AUC)
gdsc_data$Color_Nonsilent <- ifelse(gdsc_data$Mutation_Status_Nonsilent == "Other", "paleturquoise3", "turquoise4")
gdsc_data$Color_Nonsilent <- factor(gdsc_data$Color_Nonsilent)
gdsc_data <- filter(gdsc_data, !is.na(gdsc_data$AUC))
saveRDS(gdsc_data, "./data_munging/rds/ctrp_data_18Q3.rds", compress = "xz")
```

```{r}
ccle_data <- readRDS("./data_munging/rds/ccle_data_18Q3.rds")
ctrp_data <- readRDS("./data_munging/rds/ctrp_data_18Q3.rds")
gdsc_data <- readRDS("./data_munging/rds/gdsc_data_18Q3.rds")
```

```{r}
test_gdsc <- gdsc_data %>% group_by(Drug, Hugo_Symbol, Mutation_Status_Nonsilent) %>% tally()
```

# Wilcoxon tests: All genes

@@

Filter for point mutations:
```{r}
crispr_data_ptmuts <- filter(crispr_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
crispr_data_ptmuts$Keep_PtMut_Tests <- ifelse(crispr_data_ptmuts$Reference_Allele_Length == 1 | is.na(crispr_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(crispr_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
crispr_data_ptmuts <- filter(crispr_data_ptmuts, Keep_PtMut_Tests == TRUE)

@@

# G2P project

@@

@@

### CRISPR: Gene

```{r, eval=FALSE}
crispr_signif_lvlA <- compare_means(Score ~ Mutation_Status_Nonsilent, group.by = c("Hugo_Symbol"), data = crispr_data_lvlA, method = "wilcox.test", p.adjust.method = "BH")
crispr_signif_lvlA <- adj_signif(crispr_signif_lvlA)
crispr_signif_lvlA <- crispr_signif_lvlA[order(crispr_signif_lvlA$p),]
saveRDS(crispr_signif_lvlA, "./data_munging/rds/crispr_signif_lvlA_gene.rds")
```

@@

crispr_signif_genes_lvlA <- as.character(crispr_signif_lvlA$Hugo_Symbol[0:20])

knitr::kable(filter(crispr_signif_lvlA, p < 0.1)[, c("Hugo_Symbol", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "Wilcoxon test results for Level A point mutations, p < 0.01 (BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

@@

```{r fig.height=7, fig.width=12}
lvlA_genes_signif <- filter(crispr_data_ptmuts, Hugo_Symbol %in% crispr_signif_genes_lvlA)
lvlA_genes_signif <- merge(lvlA_genes_signif, crispr_signif_lvlA, by = "Hugo_Symbol")
lvlA_genes_signif$Hugo_Symbol <- ordered(lvlA_genes_signif$Hugo_Symbol, levels = crispr_signif_genes_lvlA)
lvlA_genes_signif_color <- as.character(lvlA_genes_signif$Color_Nonsilent)
names(lvlA_genes_signif_color) <- lvlA_genes_signif$Mutation_Status_Nonsilent
lvlA_genes_signif <- merge(lvlA_genes_signif, lvlA_genes_signif %>% group_by(Hugo_Symbol) %>% summarize(Text_y = max(Score)), by = "Hugo_Symbol")

lvlA_signif_genes_plot <- ggplot(data = lvlA_genes_signif, aes(x = Mutation_Status_Nonsilent, y = Score)) +
  facet_wrap(~ Hugo_Symbol, nrow = 2, scales = "free_y") +
  geom_jitter(aes(color = Mutation_Status_Nonsilent), alpha = 0.5, width = 0.3) +
  geom_boxplot(color = "black", fill = NA, outlier.shape = NA) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  scale_color_manual(values = lvlA_genes_signif_color) + 
  geom_text(data = lvlA_genes_text, mapping = aes(x = 1.5, y = Text_y, label = p.signif.adj), nudge_y = 0.2) +
  theme(legend.position = "top") +
  labs(y = "CERES Score",
       x = "Lineage",
       color = "Mutation Status",
       title = "CERES scores for select Level A G2P genes",
       subtitle = "Genes with adjusted p-values < 0.3 are shown.\nBH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001\nGenes are sorted by increasing p-value.")
lvlA_signif_genes_plot

# ggsave("./plots_18Q3/manuscript/crispr_g2p_lvlA_signif_genes_plot.png", lvlA_signif_genes_plot, width = 12, height = 7, units = "in")
```

### CRISPR: Gene and lineage

@@

```{r fig.height=15, fig.width=15}
lvlA_signif <- filter(crispr_data_ptmuts, Hugo_Symbol %in% crispr_signif_genes_lvlA)
lvlA_signif <- merge(lvlA_signif, crispr_signif_lvlA_lineage, by = c("Hugo_Symbol", "group_general_lineage_name"))
lvlA_signif$Hugo_Symbol <- ordered(lvlA_signif$Hugo_Symbol, levels = crispr_signif_genes_lvlA)
lvlA_signif_color <- as.character(lvlA_signif$Color_Nonsilent)
names(lvlA_signif_color) <- lvlA_signif$Mutation_Status_Nonsilent
lvlA_signif <- merge(lvlA_signif, lvlA_signif %>% group_by(Hugo_Symbol) %>% summarize(Text_y = max(Score)), by = "Hugo_Symbol")

lvlA_signif_lineages_plot <- ggplot(data = lvlA_signif, aes(x = group_general_lineage_name, y = Score)) +
  facet_wrap(~ Hugo_Symbol, nrow = 4, scales = "free_y") +
  geom_point(aes(color = Mutation_Status_Nonsilent), alpha = 0.5) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  scale_color_manual(values = lvlA_signif_color) +
  geom_text(data = lvlA_text, mapping = aes(x = group_general_lineage_name, y = Text_y, label = p.signif.adj), nudge_y = 0.1) +
  theme(legend.position = "top", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1, size = 10)) +
  labs(y = "CERES Score",
       x = "Lineage",
       color = "Mutation Status",
       title = "CERES score by cell line lineage for significant Level A G2P genes",
       subtitle = "Genes with adjusted p-values < 0.3 are shown.\nBH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001\nGenes are sorted by increasing lineage-agnostic p-value.")
lvlA_signif_lineages_plot

```{r}
lvlA_signif_summ <- lvlA_signif %>% group_by(Hugo_Symbol, group_general_lineage_name, Mutation_Status_Nonsilent) %>% tally()
# write.table(lvlA_signif_summ, file = "~/Desktop/lvlA_signif_summ_20181005.tsv", quote = FALSE, sep = "\t")

knitr::kable(lvlA_signif_summ, caption = "Mutant/Other counts for gene-lineage combinations") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

### CCLE: Gene and drug

```{r}
ccle_data_ptmuts <- filter(ccle_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
ccle_data_ptmuts$Keep_PtMut_Tests <- ifelse(ccle_data_ptmuts$Reference_Allele_Length == 1 | is.na(ccle_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(ccle_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
ccle_data_ptmuts <- filter(ccle_data_ptmuts, Keep_PtMut_Tests == TRUE)
ccle_data_lvlA <- filter(ccle_data_ptmuts, Hugo_Symbol %in% genes_lvlA)
```

```{r, eval=FALSE}
ccle_signif_lvlA <- compare_means(AUC ~ Mutation_Status_Nonsilent, group.by = c("Drug", "Hugo_Symbol"), data = ccle_data_lvlA, method = "wilcox.test", p.adjust.method = "BH")
ccle_signif_lvlA <- adj_signif(ccle_signif_lvlA)
ccle_signif_lvlA <- ccle_signif_lvlA[order(ccle_signif_lvlA$p),]
saveRDS(ccle_signif_lvlA, "./data_munging/rds/ccle_signif_lvlA_gene.rds")
```

```{r, eval=TRUE}
ccle_signif_lvlA <- readRDS("./data_munging/rds/ccle_signif_lvlA_gene.rds")

# ccle_signif_genes_lvlA <- as.character(ccle_signif_lvlA$Hugo_Symbol[0:20])

knitr::kable(ccle_signif_lvlA[, c("Hugo_Symbol", "Drug", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CCLE: Wilcoxon test results for Level A point mutations, p < 0.01 (BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

```{r fig.height=4, fig.width=8}
wilcox_ccle_gene_lvlA_plot <- ggplot(data = ccle_signif_lvlA) +
  geom_histogram(aes(x = p, fill = "chartreuse4"), breaks = seq(0, 1, by = 0.025), color = "black", alpha = 0.7) +
  geom_histogram(aes(x = p.adj, fill = "darkslategray3"), breaks = seq(0, 1, by = 0.025), color = "black", alpha = 0.7) +
  scale_x_continuous(breaks = seq(0, 1, by = 0.05), labels = seq(0, 1, by = 0.05)) +
  scale_fill_manual(name = "P-values", values = c("chartreuse4" = "chartreuse4", "darkslategray3" = "darkslategray3"), labels = c("Unadjusted", "BH-adjusted")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = c(0.1, 0.85)) +
  labs(x = "BH-adjusted p-values", y = "Frequency")
wilcox_ccle_gene_lvlA_plot
```

### CTRP: Gene and drug
